### 2015-04-06 cug_paper/02_design.tex
Scott's design edits

The daemon maintains a pair of lists.  One list is the blocks of GPU memory that have been assigned to the ranks of the main application, and which the ranks are presumably copying data into.  The second list is for all the blocks of memory for which the `write complete' message (step e) has been received.  When the daemon thread that handles CCI messages receives a `write done' message for a block, it moves that block from the first list to the second list.  It then notifies the background write thread(s) that the block is now ready to be written to the filesystem.

@@

Note that each write thread in the daemon has its own buffer in system memory.  Each of these buffers is deliberately kept fairly small - 16MB - in order to conserve resources.  Testing done on Titan shows that 16MB block sizes are large enough for \texttt{cudaMemcpy()} to give good performance.\cite{ac_guide} Larger buffers would simply use up memory without improving performance noticeably. (See Figure \ref{fig:transfer_bw}.\footnote{Figure \ref{fig:transfer_bw} differentiates between 'pinned' memory and 'pageable' memory.  Pageable memory is allocated using regular \texttt{malloc()} or \texttt{new} calls.  Pinned memory is allocated using \texttt{cudaMallocHost()} or \texttt{cudaHostAlloc()}.\cite{cuda_pinned_memory}  This work uses pinned memory exclusively.})

@@

There are a few peculiarities of the Cray environment that need to be taken in to account in order for this software to work properly.  The first peculiarity has to do with limiting access to the GPU and associated hardware.  The CUDA runtime can be configured to limit access to the GPU to a single thread, to multiple threads from a single process or to multiple threads from multiple processes.  On Titan, if no special options are given to qsub, the CUDA runtime will be configured for a single thread.  Since software described in this paper requires access to the GPU from multiple processes, users must switch to the appropriate compute mode.  This is done by passing the flag "-l feature=gpudefault" to qsub.\footnote{It may seem strange to have to explicitly switch to a mode called `gpudefault'.  The naming scheme comes from the CUDA documentation.  On a standard CUDA install, the GPU would in fact default to allowing access from multiple processes.  Titan is configured differently, however.  Thus the need to explicitly tell qsub to switch the GPU's to the `default' mode.}

The second peculiarity has to do with actually starting the daemon process.  Normally, executables are started on the compute nodes by aprun.  However, aprun is designed to start a single executable, not to start two different executables.  Even in MPMD mode, it will not start two different executables on the same node.  The solution is to have one rank on each node call \texttt{fork()} and then \texttt{execve()}.  Since MPI purposely hides the compute topology from the application, deciding exactly which ranks call fork() takes a little more work.  Specifically, all ranks will attempt to open a file using the \texttt{O\_CREAT} and 
\texttt{O\_EXCL} flags.  The name of the file is taken from the compute node's host name.  This combination of flags and filename ensures that exactly one rank on each node successfully opens the file.  The ranks that succeed will then start the daemon processes.

>>>>

The daemon maintains a pair of lists.  One list is the blocks of GPU memory that have been assigned to the ranks of the main application, and which the ranks are presumably copying data into.  The second list is for all the blocks of memory for which the `write complete' message (step e) has been received.  When the daemon thread that handles CCI messages receives a `write done' message for a block, it moves that block from the first list to the second list.  It then notifies the background write thread(s) that the block is now ready to be written to the filesystem. All daemon threads are pinned to unused integer cores.

@@

Note that each write thread in the daemon has its own buffer in system memory.  Each of these buffers is deliberately kept fairly small, 16MB in this test, in order to conserve resources.  Testing done on Titan shows that 16MB block sizes are large enough for \texttt{cudaMemcpy()} to give good performance.\cite{ac_guide} Larger buffers would simply use up memory without improving performance noticeably. (See Figure \ref{fig:transfer_bw}.\footnote{Figure \ref{fig:transfer_bw} differentiates between `pinned' memory and `pageable' memory.  Pageable memory is allocated using regular \texttt{malloc()} or \texttt{new} calls.  Pinned memory is allocated using \texttt{cudaMallocHost()} or \texttt{cudaHostAlloc()}.\cite{cuda_pinned_memory}  This work uses pinned memory exclusively.})

@@

There are a few peculiarities of the Cray environment that need to be taken in to account in order for this software to work properly.  The first peculiarity has to do with limiting access to the GPU and associated hardware.  The CUDA runtime can be configured to limit access to the GPU to a single thread, to multiple threads from a single process, or to multiple threads from multiple processes.  On Titan, if no special options are given to qsub, the CUDA runtime will be configured for a single thread.  Since software described in this paper requires access to the GPU from multiple processes, users must switch to the appropriate compute mode.  This is done by passing the flag ``l feature=gpudefault'' to qsub.\footnote{It may seem strange to have to explicitly switch to a mode called `gpudefault'.  The naming scheme comes from the CUDA documentation.  On a standard CUDA install, the GPU would in fact default to allowing access from multiple processes.  Titan is configured differently, however.  Thus the need to explicitly tell qsub to switch the GPU's to the `default' mode.}

The second peculiarity has to do with actually starting the daemon process.  Normally, executables are started on the compute nodes by \emph{aprun}, Cray's replacement for mpirun.  by default, aprun is designed to start a single executable in single-program-multiple-data (SPMD) mode, not to start two different executables.  Aprun does support multiple-program-multiple-data (MPMD) mode, but it will not start two different executables on the same node (i.e. each program executes on non-overlapping subsets of nodes).  The solution is to have one rank on each node call \texttt{fork()} and then \texttt{execve()}.  Since MPI purposely hides the compute topology from the application, deciding exactly which ranks call fork() takes a little more work.  Specifically, all ranks will attempt to open a file using the \texttt{O\_CREAT} and 
\texttt{O\_EXCL} flags.  The name of the file is taken from the compute node's host name.  This combination of flags and filename ensures that exactly one rank on each node successfully opens the file.  The ranks that succeed will then start the daemon processes.
